---
title: "A3: Incarceration"
output: html_document
---

### Introduction

Investigate the population of imprisoned persons per state and county which
highlights the locations in the country with high incarceration. Additionally, 
we compare the proportion of incarcerated Black Americans to white Americans.
This emphasizes the gross disproportion of incarcerated individuals and invokes
a larger question about racial discrimination in the United States.

### Summary Information
```{r summary, echo=FALSE, message=FALSE}
library(dplyr)
source("summary.R")
options(scipen = 999)
```

Through this project, it is found that as of 2018 **`r county`** with a total
population of **`r population_2018` people**. This county had the most imprisoned 
citizens with **`r average` as the average** number of incarnated people. 
We also know that on average Black Americans tend to have higher rates of 
incarceration than white Americans. In fact, the largest proportion for Black 
Americans was **`r proportion$black` in `r proportion$year`**. Finally, we 
calculated some data about the sate with the largest imprisonment rate which was 
**`r state` at `r pop`** prisoners.


### The Dataset

This particular dataset provides information about prison populations by county 
and state within the United States. The data was collected by the Vera Institute 
to expose patterns of inequality in incarceration data. The data was collected 
by compiling information from the Bureau of Justice Statistics (BJS) data 
collections. There are 153811 rows and 36 columns in the dataset which is quite 
large considering that this is only a portion of the real dataset. There are 
definitely some questions of power that need to be addressed when working with 
data that can be sensitive. It is the programmers responsibility to use this 
data to contribute to the general knowledge about this population. Data should 
not be used to push a personal narrative or agenda especially one that is harmful 
to others. Finally, it is important to note that there are some possible 
limitations with this data such as the number of missing data from the table 
musically for certain race categories. Without this data it is difficult to 
evaluate the accuracy of the provided visualizations and data summaries. This is 
not to say that this is not a credible source of information but rather 
something that has faults and should not always be taken at face value. Context 
is as important as the numerical data itself. 


### Trends Over Time Chart

This visualization displays the total prison population for the top 5 largest 
counties (in 2018) over the course of 5 years. Since this database lacks
information for prison population past 2016, it seemed more interesting to only
include valid values. This chart is included because it encompasses the most
recent data as described by the data frame. An interesting pattern that emerged
was that Harris County, GA surpassed Maricopa County, AZ in prison population
after 2014. Note that many of these counties are densely populated which 
suggests higher rates of diversity among communities. 

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("trends_over_time.R")}
``````


### Variable Comparison Chart

The following chart compares the proportion of incarcerated people to the total
population of the corresponding racial group ages 15 to 64. It is essential to
include this chart because it shows the disparities between the incarceration rate
for Black Americans versus white Americans. While Black Americans are consistently 
disproportionately imprisoned in comparison to their white counterparts there 
is a noticeable spike in the early 90s and 2000s. There could be many reasons
for this however, I believe that this was due to the influx in social justice
movements.

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("var_compare.R")}
```


### Map

This map displays the number of imprisoned persons by state. The legend to the
right of the map denotes the color to the corresponding population number. This
is important to include because it highlights visual patterns in the US. For
example, an important pattern is ...

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The last chart that you will create and include will show how a variable is distributed geographically. Again, think carefully about what such a comparison means and what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Your map needs a title
- Your color scale needs a legend with a clear label
- Use a map based coordinate system to set the aspect ratio of your map

